---
title: "Halloween Mini-project"
author: "Julia Napoli"
date: "10/29/2021"
output: html_document
---

## Importing candy data

```{r}
candy_file <- "candy-data.csv"
candy = read.csv(candy_file, row.names=1)
head(candy)
```

> Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```

> Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy$fruity)
```

## What is your favorite candy?

> Q3. What is your favorite candy? What's its winpercent value?

```{r}
rownames(candy)
candy["ReeseÕs Peanut Butter cup",]$winpercent
```
> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```


Let's look at the skimr function ...
```{r}
#install.packages("skimr")
library("skimr")
skim(candy)
```
> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Yes, the variable winpercent seems to definitely be on a different scale than the others, given its numbers range between 0-100, as opposed to 0-1.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

```{r}
candy$chocolate
```

Zero and one in this column represent whether or not the candy in question is or is not a chocolate-flavored candy.

> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```

> Q9. Is the distribution of winpercent values symmetrical?

I would say, its relatively symmetrical, but there's a minor skew right.

> Q10. Is the center of the distribution above or below 50%?

Below 50%

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
candy[as.logical(candy$chocolate),]
choc_percent <-candy$winpercent[as.logical(candy$chocolate)]
mean(candy$winpercent[as.logical(candy$chocolate)])

candy[as.logical(candy$fruity),]
fruity_percent <- candy$winpercent[as.logical(candy$fruity)]
mean(candy$winpercent[as.logical(candy$fruity)])
```


> Q12. Is this difference statistically significant?

```{r}
t.test(choc_percent,fruity_percent)
```
Yes, this is statistically significant because p<.05!


## Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

```{r}
rownames(head(candy[order(candy$winpercent),],n=5))
```

> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
rownames(head(candy[order(candy$winpercent,decreasing = TRUE),],n=5))
```


> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)
ggplot(candy) + aes(winpercent,rownames(candy)) + geom_col()
```

> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)
ggplot(candy) + aes(winpercent,reorder(rownames(candy),winpercent)) + geom_col()
```
Let's add some useful color

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"

ggplot(candy) + aes(winpercent,reorder(rownames(candy),winpercent)) + geom_col(fill=my_cols)
```
> Q17. What is the worst ranked chocolate candy?

Boston Baked Beans

> Q18. What is the best ranked fruity candy?

Starburst

## Taking a look at pricepoint

```{r}
library(ggrepel)
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

```{r}
ord <- order(candy$winpercent,decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Reeses Miniatures is the best bang for your buck.

> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Nik L Nips are the most expensive and least popular.

```{r}
my_cols[as.logical(candy$fruity)] <- "red"
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols)
```
```{r}
gsub("Õ","'",rownames(candy))
```

## Correlation Structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity & chocolate
pluribus & bar

> Q23. Similarly, what two variables are most positively correlated?

bar & chocolate
winpercent & chocolate

## Principal Component Analysis

```{r}
pca <- prcomp(candy,scale = TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2])
plot(pca$x[,1:2], col=my_cols, pch=16)
```

Make a new data-frame with PCA results and candy data.

```{r}
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)
p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```


```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Yes! This does make sense. Fruity candies are often hard and sold in baggies with multiple candies, i.e. starbursts, skittles, nerds, mike & ike, gobstoppers, etc.
