---
title: "Class 18 Worksheet"
author: "Julia Napoli"
date: "12/1/2021"
output:
  pdf_document: default
  html_document: default
---

# Section 1. Proportion of G/G in a population

Downloaded a CSV file with desired data. Now we'll read this CSV file.

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)
```

```{r}
table(mxl$Genotype..forward.strand.) / nrow(mxl) * 100
```

## Section 4: Population Analysis

> Q13. Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

How many samples do we have?

```{r}
expr <- read.table("worksheet18file.txt")
head(expr)
```

```{r}
nrow(expr)
```

Let's determine the sample size for each genotype.

```{r}
sample_size <- table(expr$geno)

sample_size
sum(sample_size)
```

Now let's find the median expression levels for each of these genotypes.

```{r}
library(dplyr)

AA_rows <- filter(expr, expr$geno == "A/A")
summary(AA_rows$exp)

AG_rows <- filter(expr, expr$geno == "A/G")
summary(AG_rows$exp)

GG_rows <- filter(expr, expr$geno == "G/G")
summary(GG_rows$exp)
```

> Q14. Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3?

Let's make a boxplot with this data.

```{r}
library(ggplot2)
```

```{r}
ggplot(expr) + aes(x=geno, y=exp, fill=geno) +
  geom_boxplot(notch = TRUE)
```

You can infer that the expression value between the A/A and G/G genotypes are stastistically significantly different, since their IQRs are entirely visually distinct from one another.


Yes, the SNP does affect the expression of ORMDL3! The G/G phenotype is associated with having reduced expression of this gene in comparison to the other genotypes.



