---
title: "Class 17 Worksheet"
author: "Julia Napoli"
date: "11/26/2021"
output: 
  pdf_document:
    latex_engine: xelatex
---

## Covid Vaccination Rates Mini-Project

First, let's import the data.

```{r}
# Import vaccination data
vax <- read.csv("vax_data.csv")
head(vax)
```

> Q1. What column details the total number of people fully vaccinated? 

the persons_fully_vaccinated column

> Q2. What column details the Zip code tabulation area? 

the zip_code_tabulation_area column

> Q3. What is the earliest date in this dataset? 

```{r}
head(sort(vax$as_of_date))
```

5-01-2021.


> Q4. What is the latest date in this dataset? 

```{r}
head(sort(vax$as_of_date, decreasing = TRUE))
```

11-23-2021.

Let's get a quick overview of the dataset.

```{r}
skimr::skim(vax)
```

> Q5. How many numeric columns are in this dataset? 

Nine.

> Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

```{r}
sum(is.na(vax$persons_fully_vaccinated))
```


> Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
round(( sum(is.na(vax$persons_fully_vaccinated)) / nrow(vax) )* 100,2)
```

> Q8. [Optional]: Why might this data be missing?

"Information redacted in accordance with CA state privacy requirement..."

## Working with Dates

```{r}
# install.packages("lubridate")
library(lubridate)
```

```{r}
today()
```

```{r}
# today() - vax$as_of_date[1]
vax$as_of_date <- ymd(vax$as_of_date)
```

```{r}
today() - vax$as_of_date[1]
```

```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q9. How many days have passed since the last update of the dataset? 

```{r}
today() - vax$as_of_date[nrow(vax)]
```

> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 

```{r}
length(unique(vax$as_of_date))
```

## Working with ZIP Codes

```{r}
# install.packages('zipcodeR')
library(zipcodeR)
```


```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```

## Focus on the San Diego area

```{r}
# Subset to San Diego county only areas
sd <- vax[vax$county == "San Diego",]
head(sd)
nrow(sd)
```

> Q11. How many distinct zip codes are listed for San Diego County? 

```{r}
length(unique(sd$zip_code_tabulation_area))
```


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd[which.max(sd$age12_plus_population),]
```

zip code 92154

```{r}
library(dplyr)
sd.11.09.21 <- filter(vax, county == "San Diego" & as_of_date == "2021-11-09")
head(sd.11.09.21)
```


> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2021-11-09”? 

```{r}
mean(sd.11.09.21$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2021-11-09”?

```{r}
hist(sd.11.09.21$percent_of_population_fully_vaccinated)
```

## Focus on UCSD/La Jolla

UC San Diego resides in the 92037 ZIP code area and is listed with an age 5+ population size of 36,144.

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```

> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
library(ggplot2)

ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

## Comparing to similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2021-11-16")

head(vax.36)
```


> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

```{r}
red_line <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
library(ggplot2)

ggplot(ucsd) +
  aes(ucsd$as_of_date,
      ucsd$percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) + geom_hline(yintercept = red_line, linetype = 'dashed', color = 'red') + 
  ylim(c(0,1)) +
  labs(x = "Date", y="Percent Vaccinated")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2021-11-16”?

```{r}
fivenum(vax.36$percent_of_population_fully_vaccinated)
mean(vax.36$percent_of_population_fully_vaccinated, na.rm = TRUE)
```


> Q18. Using ggplot generate a histogram of this data.

```{r}
hist(vax.36$percent_of_population_fully_vaccinated)

# ggplot(vax.36, aes(x = percent_of_population_fully_vaccinated)) +
  # geom_bar() +
  # labs(x = "Percent Vaccinated", y="Count")
```

> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2021-11-16") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```

zip code 92109 is above, but zip code 92040 is below

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)
head(vax.36.all)
```

```{r}
ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color= "blue") +
  labs(x= "Date", y= "Percent Vaccinated",
       title="Vaccination rate across California",
       subtitle= "Only areas with a population above 36k are shown.") +
  geom_hline(yintercept = red_line, linetype= "dashed")
```


> Q21. How do you feel about traveling for Thanksgiving and meeting for in-person class next Week?

I'm excited to be home for Thanksgiving. Either in-class or virtual next week works for me.


