---
title: "Class 06 Worksheet"
author: "Julia Napoli"
date: "10/15/2021"
output: pdf_document
---

Here I will write a function to grade student homework. I will begin with an example of student test scores in a vector.

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
student1
```

The regular average will be returned by the 'mean()' function
```{r}
mean(student1)
```

To identify the lowest test score to be dropped; to identify the position of the lowest test score to be dropped.
```{r}
min(student1)
which.min(student1)
```

This will pull the min value by indexing the position back into the original vector.
```{r}
student1[which.min(student1)]
```

ow to pull out every value but the minimum value.
```{r}
student1
student1[-8]

student1[-which.min(student1)]
```

Take the mean of the remaining values:
```{r}
mean(student1[-which.min(student1)])
```
Let's make sure it works by testing it on another student:
```{r}
student2
mean(student2[-which.min(student2)])
# Need to do some fiddling with the NA value
mean(student2,na.rm = TRUE)
```

How do I identify NA values? Use the 'is.na()' function.
```{r}
is.na(student2)
x <- student2
x[is.na(x)] <- 0
x
mean(x)
```

Time to combine all the useful pieces of script I've written together to find the average score for all students, excluding their lowest score.
```{r}
x <- student3
x[is.na(x)] <- 0 
mean(x[-which.min(x)])
```

## Now, it's time to make the function!

I will take my working snippet and make it a function.
```{r}
grade <- function(x) {
  x[is.na(x)] <- 0 
  mean(x[-which.min(x)])
}
```

Now use the function.
```{r}
grade(student1)
grade(student2)
grade(student3)
```

Time to fully **comment the code**.

```{r}
#' Calculate average homework scores (vector form)
#' while dropping the lowest score. Missing values will be
#' treated as zero.
#' 
#' @param x Numeric vector of homework scores
#'
#' @return Average score
#' @export
#'
#' @examples
#' student <- c(100,NA,90,85)
#' grade(student)
#' 
#' 
grade <- function(x) {
  # Turn missing homework scores (NA scores) to zero
  # Missing homework scores zero
  x[is.na(x)] <- 0 
  # Exclude the lowest score homework
  mean(x[-which.min(x)])
}
```

Time to take the gradebook and **grade the whole class** of multiple students.

```{r}
url <- "https://tinyurl.com/gradeinput"
gradebook <- read.csv(url, row.names =1)
gradebook
```

```{r}
apply(gradebook,1,grade)
```

> Q2. . Using your grade() function and the supplied gradebook, who is the top scoring student overall in the gradebook?

Let's take a look at the results from Q1 to find the top-scoring student.
```{r}
results <- apply(gradebook,1,grade)
sort(results, decreasing = TRUE)
```

Alternatively, I can do...
```{r}
which.max(results)
```

> Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall?

Which homework did students all around score the lowest on? Let's do a by column analysis...and calculate a summary stat for each column of the gradebook. Let's take a look at the averages, and the medians.
```{r}
# Average homework score
hw.av <- apply(gradebook, 2,mean, na.rm = TRUE)
hw.av
which.min(hw.av)

hw.med <- apply(gradebook,2,median, na.rm = TRUE)
hw.med
which.min(hw.med)
```

Interesting. The mean and medians return different results. Let's take a look at the data visually.

```{r}
boxplot(gradebook)
```

> Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)?

```{r}
results

gradebook[is.na(gradebook)] <- 0
cor(results, gradebook$hw1)
```

Let's do it on all the homeworks...
```{r}
correlations <- apply(gradebook,2,cor,x=results)
correlations
max(correlations)
```

> Q5. Make sure you save your Rmarkdown document and can click the “Knit” button to generatea PDF foramt report without errors. Finally, submit your PDF to gradescope.

