---
title: "Class 06 Homework"
author: "Julia Napoli"
date: "10/15/2021"
output: html_document
---

Section 1B.

First, we need to install the bio3d package. I will only do this once and then I will comment it out but leave it in the script.

```{r}
#install.packages("bio3d")
```

The main problem is to fix the following code:

```{r}
# Can you improve this analysis code?
library(bio3d)

s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")

s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b

plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")
```

> Q1. What type of object is returned from the read.pdb() function?

the read.pdb() function pulls a protein sequence and its corresponding information

> Q2. What does the trim.pdb() function do?

the trim.pdb() function trims the protein sequence to where the hyperbole began previously

> Q3. What input parameter would turn off the marginal black and grey rectangles in the plots and what do they represent in this case?

the sse input controls for the black & grey rectangles in the plot; leaving the sse call blank would remove the rectangles; they represent

> Q4. What would be a better plot to compare across the different proteins?



> Q5. Which proteins are more similar to each other in their B-factor trends. How could you quantify this? HINT: try the rbind(), dist() and hclust() functions together with a resulting dendrogram plot. Look up the documentation to see what each of these functions does.

> Q6. How would you generalize the original code above to work with any set of input protein structures?

Write your own function starting from the code above that analyzes protein drug interactions by reading in any protein PDB data and outputs a plot for the specified protein.

Create a new RMarkdown document with your function code AND example output. We also suggest you include narrative text that address the rubric items below. Generate a PDF report and submit this PDF to our GradeScope site.

